---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "tools/README-"
)
```


# ContourFunctions

[![Travis-CI Build Status](https://travis-ci.org/CollinErickson/contour.svg?branch=master)](https://travis-ci.org/CollinErickson/contour)
[![CRAN_Status_Badge](https://www.r-pkg.org/badges/version/ContourFunctions)](https://cran.r-project.org/package=ContourFunctions)

This is an R package that provides simple functions for creating contour plots.

## Overview

The main functions are:

* `cf_grid`: Makes a contour plot from grid data.

* `cf_func`: Makes a contour plot for a function.

* `cf_data`: Makes a contour plot for a data set by fitting a Gaussian process model.

* `cf`: Passes arguments to `cf_function` or `cf_data` depending on whether the first argument is a function or numeric.


## Installation

```
# It can be installed like any other package
install.packages("ContourFunctions")

# Or the the development version from GitHub:
# install.packages("devtools")
devtools::install_github("CollinErickson/contour")
```

## Usage

```{r cf_grid}

library(ContourFunctions)
a <- b <- seq(-4*pi, 4*pi, len = 27)
r <- sqrt(outer(a^2, b^2, "+"))
cf_grid(a, b, cos(r^2)*exp(-r/(2*pi)))
```



```{r cf_func}
f1 <- function(r) cos(r[1]^2 + r[2]^2)*exp(-sqrt(r[1]^2 + r[2]^2)/(2*pi))
cf_func(f1, xlim = c(-4*pi, 4*pi), ylim = c(-4*pi, 4*pi))
```

```{r cf_data}
set.seed(0)
x <- runif(20)
y <- runif(20)
z <- exp(-(x-.5)^2-5*(y-.5)^2)
cf_data(x,y,z)
```

